---
title: "Assignment 2"
author: 'Siddharth Harsukh Pansuria | CWID: 20005837'
date: "2/21/2022"
output:
  pdf_document: default
  html_document: default
---

### Clearing environmental variables
``` {r}
rm(list = ls())
```

### Loading breast-cancer-wisconsin.csv and replacing '?' with NA
```{r}
data <- read.csv("breast-cancer-wisconsin.csv", na.strings = '?')
```

### Viewing the top 10 rows of the loaded data
```{r}
head(data, n = 10)
```

### Deleting rows with missing values
```{r}
data <- na.omit(data)
```

### Removing the first column since it only contains ID
```{r}
data <- data[-1]
```

### Setting column 'Class' as a factor of benign and malignant
```{r}
data$Class <-factor(data$Class,levels = c(2,4), labels = c("benign","malignant"))
```

### Viewing the top 10 rows of the loaded data
```{r}
head(data, n = 10)
```

### Creating a function for min max normalization
```{r}
min_max_norm <-function(x,minX,maxX) 
{
  ans <- (x - minX) / (maxX - minX)
  return(ans) 
}
```

### Getting number of columns with the 'Class' column
```{r}
target <- ncol(data)-1
```

### Normalizing the data
```{r}
normalized_data <- data.frame(matrix(NA,nrow = nrow(data), ncol = ncol(data)))
colnames(normalized_data) <- c(colnames(data))

for (i in c(1:target)) {
  normalized_data[,i] <- min_max_norm(data[,i], min(data[,i]), max(data[,i]))
}

normalized_data[,ncol(data)] <- data[,ncol(data)]

```

### Viewing the top 10 rows of the loaded data
```{r}
head(normalized_data, n = 10)
```

### Splitting the dataset into training and test data
```{r}
id <- sort(sample(nrow(normalized_data),as.integer(.70*nrow(normalized_data))))

train_data <- normalized_data[id,]
test_data <- normalized_data[-id,]
```

### Installing knn
```{r}
install.packages("kknn",repos = "http://cran.us.r-project.org")
library(kknn)
```
### Implementing the KNN model for k = 3
```{r}
  prediction_k3 <- kknn(formula=Class~., train_data, test_data[,-11], k=3, kernel ="rectangular")
  fit_k3 <- fitted(prediction_k3)
```

### Implementing the KNN model for k = 5
```{r}
  prediction_k5 <- kknn(formula=Class~., train_data, test_data[,-11], k=5, kernel ="rectangular")
  fit_k5 <- fitted(prediction_k5)
```

### Implementing the KNN model for k = 10
```{r}
  prediction_k10 <- kknn(formula=Class~., train_data, test_data[,-11], k=10, kernel ="rectangular")
  fit_k10 <- fitted(prediction_k10)
```

### Frequency table for KNN mode where k = 3
```{r}
  table(Actual=test_data$Class, Fitted=fit_k3)
```

### Frequency table for KNN mode where k = 5
```{r}
  table(Actual=test_data$Class, Fitted=fit_k5)
```

### Frequency table for KNN mode where k = 10
```{r}
  table(Actual=test_data$Class, Fitted=fit_k10)
```