---
title: "Assignment 6"
author: 'Siddharth Harsukh Pansuria | CWID: 20005837'
date: "3/26/2022"
output:
  pdf_document: default
  html_document: default
---


### Clearing environmental variables.
```{r}
rm(list = ls())
```

### Loading the breast cancer dataset
```{r}
data <- read.csv('breast-cancer-wisconsin.csv')
```

### Displaying top 10 rows of the data loaded
```{r}
head(data, n = 10)
```

### Removing first Column since it only contains ID
```{r}
data <- data[-1]
```

### Setting column 'Class' as a factor of benign and malignant
```{r}
data$Class <- factor(data$Class, levels = c(2,4), labels = c("benign", "malignant"))
```

### Displaying top 10 rows of the data loaded
```{r}
head(data, n = 10)
```

### Splitting the dataset into training and test data
```{r}
id <- sort(sample(nrow(data), as.integer(.70 * nrow(data))))
train_data <- data[id,]
test_data <- data[-id,]
```

### Installing package and implementing C5.0
```{r}
# install.packages("C50",repos="http://cran.us.r-project.org")
library('C50')
C50 <- C5.0(Class~., data = train_data)
```

### Predicting using C5.0
```{r}
predict_C50 <- predict(C50, test_data, type = "class" )
```

### Plotting the tree
```{r}
plot(C50)
```

### Frequency table for C5.0
```{r}
table(C5.0 = predict_C50, Class = test_data$Class)
```

### Calculating the error rate
### Number of wrong predictions
```{r}
wrong_preds_C50 <- sum(predict_C50 != test_data$Class)
```

### Error Rate in C5.0 predictions
```{r}
error_rate_C50 <- wrong_preds_C50/length(predict_C50)
print(paste("Error Rate:" , error_rate_C50))
```

### Installing package and implementing Random Forest 
```{r}
# install.packages('randomForest', repos = "http://cran.us.r-project.org")
library(randomForest)
randomF <- randomForest(Class~., data = train_data, importance = TRUE, ntree = 1000)
```

### Implementing Random Forest methodology
```{r}
predict_randomF <- predict(randomF, test_data)  
```

### Identifying and plotting only the important features
```{r}  
importance(randomF)
varImpPlot(randomF, main = "Chart of variable importance")
```

### Frequency table for Random Forest
```{r}
table(RandomForest = predict_randomF, Class = test_data$Class)
```

### Calculating the error rate
### Number of wrong predictions
```{r}
wrong_preds_RF <- sum(predict_randomF != test_data$Class)
```

### Error Rate in Random Forest predictions
```{r}
error_rate_RF <- wrong_preds_RF/length(predict_randomF)
print(paste("Error Rate:" , error_rate_RF))
```