---
title: "Assignment 5"
author: 'Siddharth Harsukh Pansuria | CWID: 20005837'
date: "3/19/2022"
output:
  pdf_document: default
  html_document: default
---

### Clearing environmental variables.
```{r}
rm(list = ls())
```

### Loading the breast cancer dataset and replacing '?' with NA
```{r}
data <- read.csv('breast-cancer-wisconsin.csv')
```

### Displaying top 10 rows of the data loaded
```{r}
head(data, n=10)
```

### Removing first column since it only contains ID
```{r}
data <- data[-1]
```

### Setting column 'Class' as a factor of benign and malignant
```{r}
data$Class <- factor(data$Class, levels = c(2,4), labels = c("benign", "malignant"))
```

### Displaying top 10 rows of the data loaded
```{r}
head(data, n = 10)
```

### Splitting the dataset into training and test data
```{r}
id <- sort(sample(nrow(data), as.integer(.70 * nrow(data))))
train_data <- data[id,]
test_data <- data[-id,]
```

### Importibng library and implementing CART
```{r}
library(rpart)
cart <- rpart(Class ~ ., data = train_data, method = "class")
```

### Predicting on the test data
```{r}
predict_CART <- predict(cart, test_data, type = "class")
```

### Frequency table for CART
```{r}
table(CART = predict_CART, Class = test_data$Class)
```
### Calculating the error rate
#### Number of wrong predictions
```{r}
wrong_preds <- sum(predict_CART != test_data$Class)
```
#### Error Rate in Naïve Bayes predictions
```{r}
error_rate <- wrong_preds/length(predict_CART)
print(paste("Error Rate:" , error_rate))
```