---
title: "Assignment 1"
author: 'Siddharth Harsukh Pansuria | CWID: 20005837'
date: "2/12/2022"
output:
  pdf_document: default
  html_document: default
---


### Clearing environmental variables.
```{r}
rm(list = ls())
```

### Loading breast-cancer-wisconsin.csv and replacing '?' with NA
```{r}
data <- read.csv("breast-cancer-wisconsin.csv", na.strings = '?')
```

### Viewing the top 10 rows of the loaded data
```{r}
head(data, n = 10)
```

### Summarizing the data
```{r}
summary(data)
```

### Total number of missing values in our dataset
```{r}
sum(is.na(data))
```

### Finding which columns have missing values
```{r}
colSums(is.na(data))
```

### Finding row number of missing values
```{r}
which(is.na(data$F6))
```

### Replacing missing values with mean of the column
```{r}
for(i in 1 : ncol(data)) {
  data[is.na(data[,i]),i] <- mean(data[,i], na.rm = TRUE)
}

# rounding up columns upto 2 decimal places for simplicity
data[,-1] <-round(data[,-1],2)

print.data.frame(data)
```

### Displaying frequency table of Class vs F6
```{r}
table(data$Class,data$F6)
```

### Displaying scatter plot F1 to F6 one pair at a time
```{r}
plot(data[, 2:7], col="red", main="Scatter Plot for F1 to F6 using plot")
pairs(data[, 2:7], col="blue", main="Scatter Plot for F1 to F6 using pairs")
```

### Displaying histogram box plot for F7 to F9
```{r}
par(mfrow = c(2, 3))
hist(data$F7, prob = TRUE, main = 'F7 Histogram', xlab = 'F7',col = "red")
hist(data$F8, prob = TRUE, main = 'F8 Histogram', xlab = 'F8',col = "green")
hist(data$F9, prob = TRUE, main = 'F9 Histogram', xlab = 'F9',col = "orange")
boxplot(data$F7,main="F7 Boxplot",col = "red")
boxplot(data$F8,main="F8 Boxplot",col = "green")
boxplot(data$F9,main="F9 Boxplot",col = "orange")
```

### Deleting all objects
```{r}
ls()
rm(list = ls())
```

### Loading breast-cancer-wisconsin.csv and replacing '?' with NA
```{r}
data <- read.csv("breast-cancer-wisconsin.csv", na.strings = '?')
```

### Deleting rows with NA values
```{r}
data = na.omit(data)
```

### Final number of rows after deletion
```{r}
nrow(data)
```

### Printing the data
```{r}
print.data.frame(data)
```