---
title: "Assignment 4"
author: 'Siddharth Harsukh Pansuria | CWID: 20005837'
date: "3/6/2022"
output:
  pdf_document: default
  html_document: default
---

### Clearing Environmental Variables
```{r}
rm(list = ls())
```

### Loading the breast cancer dataset and replacing '?' with NA
```{r}
data <- read.csv("breast-cancer-wisconsin.csv", na.strings = '?')
```

### Displaying top 10 rows of the data loaded
```{r}
head(data, n = 10)
```

### Deleting rows with missing values
```{r}
data <- na.omit(data)
```

### Removing first column since it only contains ID
```{r}
data <- data[-1]
```

### Setting column 'Class' as a factor of benign and malignant
```{r}
data$Class <- factor(data$Class, levels = c(2,4), labels = c("benign", "malignant"))
```

### Displaying top 10 rows of the data loaded
```{r}
head(data, n = 10)
```

### Creating a function for min max normalization
```{r}
min_max_norm <- function(x, minX, maxX) {
  ans <- (x - minX)/(maxX - minX)
  return (ans)
}
```

### Getting number of columns without the 'Class' column
```{r}
target <- ncol(data) - 1
```

### Normalizing the data
```{r}
normalized_data <- data.frame(matrix(NA, nrow = nrow(data), ncol = ncol(data)))
colnames(normalized_data) <- c(colnames(data))

for (i in c(1 : target)) {
  normalized_data[,i] <- min_max_norm(data[,i], min(data[,i]), max(data[,i]))
}

normalized_data[,ncol(data)] <- data[,ncol(data)]
```

### Displaying top 10 rows of the data loaded
```{r}
head(normalized_data, n = 10)
```

### Splitting the dataset into training and test data
```{r}
id <- sort(sample(nrow(normalized_data), as.integer(.70 * nrow(normalized_data))))
train_data <- normalized_data[id,]
test_data <- normalized_data[-id,]
```

### Installing package to implement Na誰ve Bayes methodology 
```{r}
install.packages("e1071", dependencies = TRUE, repos = "http://cran.us.r-project.org")
library(e1071)
```

### Implementing Na誰ve Bayes methodology
```{r}
NaiveBayes <- naiveBayes(Class ~., data = train_data)
predict_NaiveBayes <- predict(NaiveBayes, test_data)
```

### Frequency table for the Na誰ve Bayes predictions
```{r}
table(NaiveBayes = predict_NaiveBayes, Class = test_data$Class)
```

### Calculating the error rate
#### Number of wrong predictions
```{r}
wrong_preds <- sum(predict_NaiveBayes != test_data$Class)
```
#### Error Rate in Na誰ve Bayes predictions
```{r}
error_rate <- wrong_preds/length(predict_NaiveBayes)
print(paste("Error Rate:" , error_rate))
```
